---
title: 'STAT 231: R Q & A Discussion Forum'
author: 'Sean, Tamer, Chris'
date: "Sept. 3, 2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, setup, include=FALSE}
# Load packages here 
library(mdsr)   
library(babynames)
library(tidyverse)

knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small",
  message=FALSE)   # slightly smaller font for code
```

### Q11: What are the differences between facet_grid() and facet_wrap()?  Can a quantitative variable be used within either of these facet functions?

SOLUTION: Facet_wrap(~variable) will return a symmetrical matrix of plots for the number of levels of variable, while facet_grid(. ~variable) will return facets equal to the levels of variable distributed horizontally. An example of this can be seen below.

If a quantitative variable is being faceted, it will be converted to a categorical variable, and the plot will contain a facet for each distinct value.

http://www.cookbook-r.com/Graphs/Facets_(ggplot2)/

https://stackoverflow.com/questions/20457905/whats-the-difference-between-facet-wrap-and-facet-grid-in-ggplot2


```{r}
library(nasaweather)

#Using Facet Wrap
ggplot(storms, aes(x = long, y = lat)) + 
  geom_path(aes(color = name)) + 
  facet_wrap(~year) + 
  scale_color_discrete(guide="none") + 
  labs(x = "Longitude", y = "Latitude") +
  ggtitle("Tropical Storm Paths From 1995-2000 Using Facet Wrap")

#Using Facet Grid
ggplot(storms, aes(x = long, y = lat)) + 
  geom_path(aes(color = name)) + 
  facet_grid(~year) + 
  scale_color_discrete(guide="none") + 
  labs(x = "Longitude", y = "Latitude") +
  ggtitle("Tropical Storm Paths From 1995-2000 Using Facet Grid")
```


 