---
title: 'STAT 231: Problem Set 8A'
author: "Sean Wei"
date: "due by 5 PM on Monday, November 2"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

In order to most effectively digest the textbook chapter readings -- and the  new R commands each presents -- series A homework assignments are designed to encourage you to read the textbook chapters actively and in line with the textbook's Prop Tip of page 33:

"\textbf{Pro Tip}: If you want to learn how to use a particular command, we highly recommend running the example code on your own" 

A more thorough reading and light practice of the textbook chapter prior to class allows us to dive quicker and deeper into the topics and commands during class.  Furthermore, learning a programming lanugage is like learning any other language -- practice, practice, practice is the key to fluency.  By having two assignments each week, I hope to encourage practice throughout the week.  A little coding each day will take you a long way!

*Series A assignments are intended to be completed individually.*  While most of our work in this class will be collaborative, it is important each individual completes the active readings.  The problems should be straightforward based on the textbook readings, but if you have any questions, feel free to ask me!

Steps to proceed:

\begin{enumerate}
\item In RStudio, go to File > Open Project, navigate to the folder with the course-content repo, select the course-content project (course-content.Rproj), and click "Open" 
\item Pull the course-content repo (e.g. using the blue-ish down arrow in the Git tab in upper right window)
\item Copy ps8A.Rmd from the course repo to your repo (see page 6 of the GitHub Classroom Guide for Stat231 if needed)
\item Close the course-content repo project in RStudio
\item Open YOUR repo project in RStudio
\item In the ps8A.Rmd file in YOUR repo, replace "YOUR NAME HERE" with your name
\item Add in your responses, committing and pushing to YOUR repo in appropriate places along the way
\item Run "Knit PDF" 
\item Upload the pdf to Gradescope.  Don't forget to select which of your pages are associated with each problem.  \textit{You will not get credit for work on unassigned pages (e.g., if you only selected the first page but your solution spans two pages, you would lose points for any part on the second page that the grader can't see).} 
\end{enumerate}

```{r, setup, include=FALSE}
library(tidyverse)
library(mdsr)

knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```


\newpage
# 1.  k-means clustering

Section 9.1.2 walks through an example of how k-means clustering can identify genuine patterns in data -- in this case, clustering cities into continental groups merely based on city location (latitutde and longitude coordinates).  The code below functions similarly to the code provided in this section of the textbook, but uses slightly different syntax.  It also adds the centroid locations to a reproduction of Figure 9.4.

(a) Walk through the code and comments below to understand what the code is doing.  How many cities were identified as belonging to cluster 1?  What does cluster 1 represent?

> ANSWER: There are 726 cities that belong to cluster 1. Cluster 1 represents the group of cities that are the most close to a latitude coordinate of 27.43226 and a longitude coordinate of 75.14407.

```{r}
# the code in the textbook loads WorldCities, but the dataset name has since 
# been updated to world_cities 
# (world_cities is a dataset provided in the mdsr package)
data(world_cities)

# identify the 4,000 biggest cities in the world
BigCities <- world_cities %>%
  arrange(desc(population)) %>%
  head(4000) %>%
  select(latitude, longitude)

# apply the k-means algorithm (set the seed to make the results reproducible)
set.seed(15)
km_out <- kmeans(BigCities, centers = 6, nstart = 20)

# km_out is of class "kmeans"; it's a kmeans object
class(km_out)
# you can use the syntax km_out$name to refer to any of the named elements below
names(km_out)
# for example, let's see what the cluster sizes are (how many cities are in each cluster)
km_out$size
# for another example, let's see where the centroids are
km_out$centers

BigCities_km <- BigCities %>%
  mutate(cluster = as.character(km_out$cluster))

ggplot(data = BigCities_km, aes(x = longitude, y = latitude)) + 
  geom_point(aes(color = cluster), alpha = 0.5) +
  # add centroids to plot
  geom_point(data = as.data.frame(km_out$centers), shape = "X", size = 3)
```

(b) In k-means clustering, the analyst specifies the number of clusters to create.  Update the `center` argument within the `kmeans` function to identify 3 clusters instead of 6.  Create a plot like the one above, but coloring the points by these new cluster assignments.  How many cities are in cluster 1 now?  What does cluster 1 represent now?

> ANSWER: After using 3 clusters, there are now 1466 cities in cluster 1. Cluster 1 now represents all cities that are located around a latitude coordinate of 33.29923 and a longitude coordinate of 21.65145.

```{r}
# apply the k-means algorithm (set the seed to make the results reproducible)
set.seed(15)
km_out <- kmeans(BigCities, centers = 3, nstart = 20)

# cluster sizes
km_out$size

#centroids
km_out$centers

BigCities_km <- BigCities %>%
  mutate(cluster = as.character(km_out$cluster))

ggplot(data = BigCities_km, aes(x = longitude, y = latitude)) + 
  geom_point(aes(color = cluster), alpha = 0.5) +
  # add centroids to plot
  geom_point(data = as.data.frame(km_out$centers), shape = "X", size = 3)
```

(c) Lastly, update the `center` argument within the `kmeans` function to identify 15 clusters.  Create a plot like the one above, but coloring the points by these new cluster assignments.  How many cities are in cluster 1 now?  What does cluster 1 represent now?

> ANSWER: After using 15 clusters, there are now 593 cities in cluster 1. Cluster 1 represents the cities that are located near a latitude coordinate of 25.421427 and a longitude coordinate of 79.122413.

```{r}
# apply the k-means algorithm (set the seed to make the results reproducible)
set.seed(15)
km_out <- kmeans(BigCities, centers = 15, nstart = 20)

# cluster sizes
km_out$size

#centroids
km_out$centers

BigCities_km <- BigCities %>%
  mutate(cluster = as.character(km_out$cluster))

ggplot(data = BigCities_km, aes(x = longitude, y = latitude)) + 
  geom_point(aes(color = cluster), alpha = 0.5) +
  # add centroids to plot
  geom_point(data = as.data.frame(km_out$centers), shape = "X", size = 3)
```




